---
title: "DS Indy Visualization pt2 R"
author: "David Taylor"
date: "February 9, 2016"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

# Install Referenced Packages

###install.packages("ggplot2")
###install.packages("dplyr")
###install.packages("AppliedPredictiveModeling")
###install.packages("caret")
###install.packages("compute")
###install.packages("Hmisc")
###install.packages("vcd")
###install.packages("corrgram")
###install.packages("vcd")
###install.packages("qrage")
###install.packages("shiny")
###install.packageS("DT")

# Lines


```{r}
# Create Line Chart

# convert factor to numeric for convenience 
Orange$Tree <- as.numeric(Orange$Tree) 
ntrees <- max(Orange$Tree)

# get the range for the x and y axis 
xrange <- range(Orange$age) 
yrange <- range(Orange$circumference) 

# set up the plot 
plot(xrange, yrange, type="n", xlab="Age (days)",
  	ylab="Circumference (mm)" ) 
colors <- rainbow(ntrees) 
linetype <- c(1:ntrees) 
plotchar <- seq(18,18+ntrees,1)

# add lines 
for (i in 1:ntrees) { 
  tree <- subset(Orange, Tree==i) 
  lines(tree$age, tree$circumference, type="b", lwd=1.5,
    lty=linetype[i], col=colors[i], pch=plotchar[i]) 
} 

# add a title and subtitle 
title("Tree Growth", "example of line plot")

# add a legend 
legend(xrange[1], yrange[2], 1:ntrees, cex=0.8, col=colors,
  	pch=plotchar, lty=linetype, title="Tree")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Bar


## Nominal vs. Quantative Continuous

```{r}


library(ggplot2)
library(dplyr)

foobar <- aggregate(iris$Sepal.Width, list(iris$Species), mean)

foo <- rename(foobar, Species=Group.1, "Average.Sepal.Width"=x)

ggplot(foo, aes(x = Species, y=Average.Sepal.Width, fill=Species)) +  geom_bar(stat ="identity")

```


  
```{r}

foobar <- mtcars[c("cyl","mpg")]

row.names(foobar) <- NULL

foo <- aggregate(foobar, list(foobar$cyl), mean)

ggplot(foo, aes(x=cyl, y=mpg)) + geom_bar(stat="identity")


```


## Discrete vs. Discrete

```{r}
# Simple Bar Plot 


barplot(table(mtcars$cyl), main="Plot of Discrete Cylindars vs. Counts of Cars with respective Cylindars")


```


## Discrete vs. Continuous

```{r}

foobar <- mtcars[c("mpg", "cyl")]

row.names(foobar) = NULL

ggplot(foobar, aes(y = mpg, x = cyl, fill = cyl)) +  
  geom_bar(position = "dodge",stat = "identity") 

```



# Scatter

## Continuous vs. Continuous

```{r}

ggplot(mtcars, aes(disp, mpg)) + geom_point(aes(color=cyl))

```

## Matrix

```{r}

library(AppliedPredictiveModeling)
transparentTheme(trans = .4)

library(caret)
featurePlot(x = iris[, 1:4],
            y = iris$Species,
            plot = "pairs",
            ## Add a key at the top
            auto.key = list(columns = 3))

```

# Gantt

## Continuous vs. Nominal 

```{r}

foobar <- aggregate(iris$Sepal.Width, list(iris$Species), mean)

foo <- rename(foobar, Species=Group.1, "Average.Sepal.Width"=x)

# NOTE THAT YOU NEED TO FLIP WHAT WILL BE HORIZONTAL and VERTICAL AXIS SERIES when YOU USE coord_flip()
ggplot(foo, aes(y=Average.Sepal.Width, x=Species)) + geom_bar(stat="identity") + coord_flip()

```


## Continuous vs. Discrete 


```{r}

foobar <- mtcars[c("mpg", "cyl")]

foo <- aggregate(foobar, list(foobar$cyl), mean)

show(foo)

ggplot(foobar, aes(x=mpg, y=cyl)) + geom_bar(stat="identity") + coord_flip()

```


#   TABLE PLOT
```{r}

library(DT)
datatable(iris)




```

# Mosaics
```{r}
# Mosaic Plot Example
library(vcd)
mosaic(HairEyeColor, shade=TRUE, legend=TRUE)
```


#Correlograms
```{r}
library(corrgram)
corrgram(mtcars, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Car Milage Data in PC2/PC1 Order")
```


# Second Correlogram Example
```{r}
library(corrgram)
corrgram(mtcars, order=TRUE, lower.panel=panel.ellipse,
  upper.panel=panel.pts, text.panel=panel.txt,
  diag.panel=panel.minmax, 
 	main="Car Milage Data in PC2/PC1 Order")
```



# Probability Plots
# Display the Student's t distributions with various
# degrees of freedom and compare to the normal distribution
```{r}
x <- seq(-4, 4, length=100)
hx <- dnorm(x)

degf <- c(1, 3, 8, 30)
colors <- c("red", "blue", "darkgreen", "gold", "black")
labels <- c("df=1", "df=3", "df=8", "df=30", "normal")

plot(x, hx, type="l", lty=2, xlab="x value",
  ylab="Density", main="Comparison of t Distributions")

for (i in 1:4){
  lines(x, dt(x,degf[i]), lwd=2, col=colors[i])
}

legend("topright", inset=.05, title="Distributions",
  labels, lwd=2, lty=c(1, 1, 1, 1, 2), col=colors)
```  


# Children's IQ scores are normally distributed with a
# mean of 100 and a standard deviation of 15. What
# proportion of children are expected to have an IQ between
# 80 and 120?
```{r}
mean=100; sd=15
lb=80; ub=120

x <- seq(-4,4,length=100)*sd + mean
hx <- dnorm(x,mean,sd)

plot(x, hx, type="n", xlab="IQ Values", ylab="",
  main="Normal Distribution", axes=FALSE)

i <- x >= lb & x <= ub
lines(x, hx)
polygon(c(lb,x[i],ub), c(0,hx[i],0), col="red") 

area <- pnorm(ub, mean, sd) - pnorm(lb, mean, sd)
result <- paste("P(",lb,"< IQ <",ub,") =",
   signif(area, digits=3))
mtext(result,3)
axis(1, at=seq(40, 160, 20), pos=0)

# Q-Q plots
```{r}
par(mfrow=c(1,2))

# create sample data 
x <- rt(100, df=3)

# normal fit 
qqnorm(x); qqline(x)

# t(3Df) fit 
qqplot(rt(1000,df=3), x, main="t(3) Q-Q Plot", 
   ylab="Sample Quantiles")
abline(0,1)


```


# Density Plots

```{r}
library(caret)
transparentTheme(trans = .9)
featurePlot(x = iris[, 1:4],
                  y = iris$Species,
                  plot = "density",
                  ## Pass in options to xyplot() to 
                  ## make it prettier
                  scales = list(x = list(relation="free"),
                                y = list(relation="free")),
                  adjust = 1.5,
                  pch = "|",
                  layout = c(4, 1),
                  auto.key = list(columns = 3))

```

# HISTORAMs

```{r}
library(ggplot2)

BMI<-rnorm(n=1000, m=24.2, sd=2.2) 

hist(BMI)

hist(BMI, freq=FALSE,
     xlab="Body Mass Index",
     main="Distribution of Body Mass Index",
     col="lightgreen",
     xlim=c(15,35),
     ylim=c(0, .20)
     )
curve(dnorm(x, mean=mean(BMI), sd=sd(BMI)), add=TRUE, col="darkblue", lwd=2)



```

```{r}

# ggplot2 examples
library(ggplot2) 
data(mtcars)

# create factors with value labels 
mtcars$gear <- factor(mtcars$gear,levels=c(3,4,5),
  	labels=c("3gears","4gears","5gears")) 
mtcars$am <- factor(mtcars$am,levels=c(0,1),
  	labels=c("Automatic","Manual")) 
mtcars$cyl <- factor(mtcars$cyl,levels=c(4,6,8),
   labels=c("4cyl","6cyl","8cyl")) 

# summary(mtcars$mpg)
# labels(mtcars)
# head(mtcars)



# Kernel density plots for mpg
# grouped by number of gears (indicated by color)
qplot(mpg, data=mtcars, geom="density", fill=gear, alpha=I(.5), 
   main="Distribution of Gas Milage", xlab="Miles Per Gallon", 
   ylab="Density")

# summary(mtcars$hp)
# summary(mtcars$gear)
# summary(mtcars$cyl)
# summary(mtcars$am)

# Scatterplot of mpg vs. hp for each combination of gears and cylinders
# in each facet, transmittion type is represented by shape and color
qplot(hp, mpg, data=mtcars, shape=am, color=am, 
   facets=gear~cyl, size=I(3),
   xlab="Horsepower", ylab="Miles per Gallon") 

# Separate regressions of mpg on weight for each number of cylinders
qplot(wt, mpg, data=mtcars, geom=c("point", "smooth"), 
#   method="lm",
#   formula=y~x, 
   color=cyl, 
   main="Regression of MPG on Weight", 
   xlab="Weight", ylab="Miles per Gallon")

# Boxplots of mpg by number of gears 
# observations (points) are overlayed and jittered
qplot(gear, mpg, data=mtcars, geom=c("boxplot", "jitter"), 
   fill=gear, main="Mileage by Gear Number",
   xlab="", ylab="Miles per Gallon")

```

# 
```{r}
data(mtcars)
mymtcars<-mtcars

mymtcars$country<- c(rep("JP",3),rep("US",4),rep("DE",7),rep("US",3),"IT",rep("JP",3),rep("US",4),"IT","DE","UK","US",rep("IT",2),"SE")

ggplot(mymtcars, aes(x=wt, y=mpg, group=country)) + geom_line(aes(colour=country))
```


# Dendrogram
```{r}
# prepare hierarchical cluster
hc = hclust(dist(mtcars))
# very simple dendrogram
plot(hc)
```




# ROC

http://www.r-bloggers.com/a-small-introduction-to-the-rocr-package/

```{r}

library(ROCR)

data(ROCR.simple)

pred <- prediction(ROCR.simple$predictions,ROCR.simple$labels)
class(pred)

slotNames(pred)

sn = slotNames(pred)
sapply(sn, function(x) length(slot(pred, x)))

sapply(sn, function(x) class(slot(pred, x)))


data(ROCR.hiv)
manypred = prediction(ROCR.hiv$hiv.nn$predictions, ROCR.hiv$hiv.nn$labels)
sapply(sn, function(x) length(slot(manypred, x)))

sapply(sn, function(x) class(slot(manypred, x)))

#performance(prediction.obj, measure, x.measure='cutoff', ...)

roc.perf = performance(pred, measure = 'tpr', x.measure = 'fpr')
plot(roc.perf)
abline(a=0, b= 1)


many.roc.perf = performance(manypred, measure = 'tpr', x.measure = 'fpr')
plot(many.roc.perf, col=1:10)
abline(a=0, b= 1)


```
 


